what will you focus on/
what makes you different from the other group?
- the focus on pipeline, modular, bla bla, could even integrate other groups work at the end

how will you ensure quality of product?
- testing, branch coverage, other?

why terra? - used by biologists, jupyter notebook like, easy to use, cloud based

docker? - ask somebody else. petar/daglar

midterm meeting prototype - will include must haves

\item The program \textbf{must} be able to use sample data residing on the same platform as the program.
\item The program \textbf{must} accept a configuration file in a machine readable format containing preset settings as well as a file path to the data sample.
\item The program \textbf{must} be able to use at least one existing strain estimation tool to compute a prediction of the relative abundancies of the strains present in the sample. \textbf{*}
\item After the strains have been estimated the program \textbf{must} offer a report with the results in a standardised machine- and human-readable format.

+ more?

prototype questions - petar

petar - strainEst on terra
konsta - general stuff
yiran - algorithm stuff. preprocessing/tools
daglar - terra

scalable?
parallel computing?
what's cool about it?
relative abundances?

do you use terra/wdl(petar) just because lucas wanted to? or did you actually consider it was best?
how did you compare it? - daglar 

google cloud costs?   - 300 dollars free credits, estimated charges so far (close to 0)
already use sickle/bowtie?  - yes, by strainest               

decrease algo priority
increase gui/ lines of code priority

